# Data descriptives

```{r setup, include=FALSE}
pacman::p_load(
  knitr, 
  scales,
  broom, 
  ggstance,
  tidyverse
)
opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2, 
  fig.width = 8,
  fig.align = 'center'
)
theme_set(
  theme_linedraw() +
    theme(
      panel.grid = element_blank()
    )
)
```

## Data

```{r data-read}
mtf <- read_rds("data/mtf.rds")
us <- read_rds("data/us.rds")
yrbs <- read_rds("data/yrbs.rds")
```

### Sample sizes and years

Sample sizes per dataset are the number of rows that include the outcome variable (there are two for US). The total sample size *in the entire manuscript* is the sum without US-emotion (including both US outcomes would duplicate its N).

```{r}
tibble(
  mtf = drop_na(mtf, Depression) %>% nrow,
  us_conduct = drop_na(us, Conduct) %>% nrow,
  us_emotion = drop_na(us, Emotion) %>% nrow,
  yrbs = drop_na(yrbs, Suicide) %>% nrow
) %>% 
  mutate(total = mtf + us_conduct + yrbs) %>% 
  mutate(across(.fns = ~comma(.))) %>% 
  kable()
```

Actual sample sizes used in analyses are the intersections of non-missing values for the two variables (outcome, technology) used in each model.

```{r}
doit <- function(data, name, x, y) {
  data <- data %>% 
    drop_na(all_of(x), all_of(y))
  data %>% 
    count(Sex) %>% 
    pivot_wider(names_from = Sex, values_from = n) %>% 
    mutate(Total = Female + Male) %>%
    mutate(
      across(
        c(Female, Male), 
        ~str_glue("{comma(.)} ({percent(./Total)})")
      )
    ) %>% 
    mutate(across(where(is.numeric), ~comma(.))) %>% 
    mutate(Years = paste(range(data$Year), collapse = " - ")) %>% 
    mutate(Dataset = name, Technology = x, Outcome = y) %>% 
    select(Dataset, Technology, Outcome, Years, Total, Female, Male)
}
bind_rows(
  doit(mtf, "MTF", "TV", "Depression"),
  doit(mtf, "MTF", "SM", "Depression"),
  doit(us, "US", "TV", "Emotion"),
  doit(us, "US", "SM", "Emotion"),
  doit(us, "US", "TV", "Conduct"),
  doit(us, "US", "SM", "Conduct"),
  doit(yrbs, "YRBS", "TV", "Suicide"),
  doit(yrbs, "YRBS", "DV", "Suicide")
) %>% 
  arrange(Dataset, Technology, Outcome) %>% 
  kable()
```

### Outcomes

```{r}
mtf %>% 
  select(starts_with("D_B_")) %>% 
  psych::alpha() %>% 
  summary
cor(
  select(mtf, starts_with("D_B_")), 
  use = "pairwise.complete.obs"
) %>% 
  round(2)
```

```{r}
sdq_con <- c("sdqe", "sdqg", "sdql", "sdqr", "sdqv")
sdq_emo <- c("sdqc", "sdqh", "sdqm", "sdqp", "sdqx")
us %>% 
  select(all_of(sdq_con)) %>% 
  psych::alpha() %>% 
  summary
us %>% 
  select(all_of(sdq_emo)) %>% 
  psych::alpha() %>% 
  summary
cor(
  select(us, all_of(sdq_con)), 
  use = "pairwise.complete.obs"
) %>% 
  round(2)
cor(
  select(us, all_of(sdq_emo)), 
  use = "pairwise.complete.obs"
) %>% 
  round(2)
```

## Figures

```{r mtf-data-fig, cache = TRUE}
mtf %>% 
  pivot_longer(c(TV, SM, Depression)) %>% 
  drop_na(value) %>% 
  add_count(name) %>% 
  mutate(name = str_glue("{name} (N = {comma(n)})")) %>% 
  ggplot(aes(Year, value)) +
  geom_blank() +
  stat_summary(fun=mean, geom="line") +
  facet_wrap("name", scales = "free_y")
```

```{r us-data-fig, cache = TRUE}
us %>% 
  pivot_longer(c(SM, TV, Emotion, Conduct)) %>% 
  drop_na(value) %>% 
  add_count(name) %>% 
  mutate(name = str_glue("{name} (N = {comma(n, accuracy = 1)})")) %>%
  ggplot(aes(Year, value)) +
  geom_blank() +
  stat_summary(fun=mean, geom="line") +
  facet_wrap("name", scales = "free_y")
```

```{r yrbs-data-fig, cache = TRUE}
yrbs %>% 
  pivot_longer(c(TV, DV, sad_lonely:suicide_3)) %>% 
  drop_na(value) %>% 
  add_count(name) %>% 
  mutate(name = str_glue("{name} (N = {comma(n)})")) %>%
  ggplot(aes(Year, value)) +
  geom_blank() +
  stat_summary(fun=mean, geom="line") +
  facet_wrap("name", scales = "free_y")
```
