# Bayesian Generalized Linear Mixed Model

```{r setup, include=FALSE}
pacman::p_load(
  knitr, 
  brms, 
  future,
  ggstance,
  patchwork,
  tidyverse
)
opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2, 
  fig.width = 6,
  fig.height = 4,
  fig.align = 'center',
  # Save more figure formats
  dev = c("png", "pdf", "tiff"),
  fig.ext = c("png", "pdf", "tiff")
)
theme_set(
  theme_linedraw(base_size = 9) +
    theme(
      panel.grid = element_blank()
    )
)
dir.create("models")
```

For each dataset and outcome, we fit a GLMM using brms, in order to do full bayesian inference. These models take a while to run. 

## Data

```{r data-load}
mtf <- read_rds("data/mtf.rds")
us <- read_rds("data/us.rds")
yrbs <- read_rds("data/yrbs.rds")
```

For US, we also isolate between- and within-person variables.

```{r data-prep-us}
# Within-person center variables (Vuorre & Bolger, 2017)
us <- bmlm::isolate(us, "pidp", "TV", which = "both", z = TRUE)
us <- bmlm::isolate(us, "pidp", "SM", which = "both", z = TRUE)
us <- us %>% select(-TV, -SM)
# We focus on between-person relations so rename those back
us <- us %>% rename(TV = TV_cb, SM = SM_cb)
```

YRBS is in long format wrt items

```{r data-prep-yrbs}
yrbs <- yrbs %>% 
  select(-suicide_4, -Suicide) %>% 
  pivot_longer(
    sad_lonely:suicide_3, 
    names_to = "item", values_to = "Suicide"
  )
```

## Models

### Estimate

- MTF: Single level model
- US: Random intercepts for participants. 85% of participants have only 1 or 2 observations, so random slopes without strong priors are not a good idea.
- YRBS: Full item-specific random effects.

```{r}
fit <- function(data, name, x, y, iter = 1000, chains = 12, cores = 12) {
  data <- drop_na(data, all_of(x), all_of(y))
  data <- mutate(data, Year = Year - 2017)
  if (name=="YRBS") {
    data <- data %>% 
      mutate(across(all_of(x), ~as.numeric(scale(.))))
    ml1 <- brm(
      str_glue("{y} ~ Sex * Year * {x} + (Sex * Year * {x} | item)"), 
      family = bernoulli("probit_approx"),
      control = list(adapt_delta = .98),
      prior = prior(student_t(9, 0, 1), class = "sd") +
        prior(normal(0, 1), class = "b"),
      inits = "0",
      data = data, chains = chains, iter = iter, cores = cores,
      file = str_glue("models/brm-{name}-{x}-{y}-ml1-probit")
    )
  } else if (name=="MTF") {
    data <- data %>% 
      mutate(across(c(all_of(x), all_of(y)), ~as.numeric(scale(.))))
    ml1 <- brm(
      str_glue("{y} ~ Sex * Year * {x}"), 
      control = list(adapt_delta = .9),
      data = data, chains = chains, iter = iter, cores = cores,
      file = str_glue("models/brm-{name}-{x}-{y}-ml1")
    )
  } else if (name=="US") {
    data <- data %>% 
      mutate(across(c(all_of(x), all_of(y)), ~as.numeric(scale(.))))
    ml1 <- brm(
      str_glue("{y} ~ Sex * Year * {x} + (1 | pidp)"), 
      control = list(adapt_delta = .9),
      data = data, chains = chains, iter = iter, cores = cores,
      file = str_glue("models/brm-{name}-{x}-{y}-ml1")
    )
  }
  tibble(
    data = name,
    Technology = x,
    Outcome = y,
    ml1 = list(ml1)
  )
}
```

```{r models-run, eval=TRUE, cache=FALSE}
x1 <- fit(yrbs, "YRBS", "TV", "Suicide")
x2 <- fit(yrbs, "YRBS", "DV", "Suicide")
x3 <- fit(mtf, "MTF", "TV", "Depression")
x4 <- fit(mtf, "MTF", "SM", "Depression")
x5 <- fit(us, "US", "SM", "Emotion")
x6 <- fit(us, "US", "SM", "Conduct")
x7 <- fit(us, "US", "TV", "Emotion")
x8 <- fit(us, "US", "TV", "Conduct")
```

```{r}
# Rename variables
fits <- bind_rows(x1,x2,x3,x4,x5,x6,x7,x8)
fits <- fits %>% 
  mutate(
    Technology = ifelse(
      Technology %in% c("SM", "DV"), 
      "Social media / device", 
      "Television"
    )
  )
```

### Results

Focus on some YRBS outcomes: plan suicide (suicide_2) on social media

```{r}
fits

fitted(fits$ml1[[2]], newdata = tibble(item = "suicide_2", Sex = "Male", Year = 0, DV = 0:1), summary = FALSE, re_formula = NULL) %>% 
  as.data.frame() %>% 
  tibble() %>% 
  mutate(delta = V2-V1) %>%
  pivot_longer(everything()) %>% 
  group_by(name) %>% 
  tidybayes::mean_qi(value)
```


```{r Parameter-figure}
probs <- c(.025, .1, .9, .975)
pars <- fits %>% 
  mutate(
    p = map(
      ml1, ~posterior_summary(., pars = "^b_", probs = probs) %>%
        as.data.frame() %>% 
        rownames_to_column("Parameter") %>% 
        tibble()
    )
  ) %>%
  unnest(p) %>% 
  select(-ml1)
b_Year <- pars %>% 
  filter(Parameter == "b_Year") %>% 
  mutate(Parameter = "Year")
b_Tech <- pars %>% 
  filter(Parameter %in% c("b_TV", "b_DV", "b_SM")) %>% 
  mutate(Parameter = "Technology")
b_Interaction <- pars %>% 
  filter(Parameter %in% c("b_Year:TV", "b_Year:DV", "b_Year:SM")) %>% 
  mutate(Parameter = "Year by Technology Interaction")
p1 <- bind_rows(b_Year, b_Tech, b_Interaction) %>% 
  mutate(Outcome = fct_rev(Outcome)) %>% 
  # Fill points based on if 95%CI includes zero
  mutate(
    Zero = ifelse(
      sign(Q2.5) == sign(Estimate) & sign(Estimate) == sign(Q97.5), 
      "*", ""
    )
  ) %>% 
  mutate(
    Parameter = factor(Parameter, levels = c("Year", "Technology", "Year by Technology Interaction"))
  ) %>% 
  ggplot(aes(Estimate, Outcome, shape = Technology, fill = Zero)) +
  scale_shape_manual(values = c(21, 22)) +
  scale_fill_manual(values = c("white", "black"), guide = FALSE) +
  scale_x_continuous(
    "Parameter estimate",
    breaks = scales::pretty_breaks(), 
    expand = expansion(.15)
  ) +
  geom_vline(xintercept = 0, lty = 2, size = .25) +
  geom_linerangeh(
    aes(xmin = Q2.5, xmax = Q97.5), size = .25,
    position = position_dodge2v(.5), show.legend = FALSE
  ) +
  geom_linerangeh(
    aes(xmin = Q10, xmax = Q90), size = .75,
    position = position_dodge2v(.5), show.legend = FALSE
  ) +
  geom_point(
    size = 2, position = position_dodge2v(.5),
  ) +
  facet_wrap("Parameter", scales = "free_x", nrow = 1) +
  theme(
    legend.position = "right",
    axis.title.y = element_blank(), 
    panel.spacing.x = unit(12, "pt")
  )

((p1 %+% 
    filter(p1$data, Parameter %in% c("Technology")) +
    theme(axis.title.x = element_blank())) /
  p1 %+% filter(p1$data, Parameter == "Year by Technology Interaction")) +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")
```


```{r, eval = FALSE}
# Get YRBS item-specific effects
tmp1 <- slice(fits, 1:2) %>% 
  mutate(
    o = map(ml1, ~rownames_to_column(as.data.frame(coef(.)$item[,,7])))
  ) %>% 
  unnest(o)

tmp <- fits %>% 
  mutate(
    p = map(
      ml1, ~posterior_summary(., pars = "b_Year:[TV|SM|DV]") %>%
        as.data.frame() %>% 
        rownames_to_column() %>% 
        tibble()
    )
  ) %>%
  unnest(p)
tmp %>%   
  ggplot(aes(Estimate, Outcome, col = Outcome)) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(
    "Linear interaction estimate (±95%CI)",
    breaks = scales::pretty_breaks(), 
    expand = expansion(.2)
  ) +
  geom_vline(xintercept = 0, size = .25, lty = 2) +
  ggstance::geom_pointrangeh(
    shape = 21, fill = "white", fatten = 3,
    aes(xmin=Q2.5, xmax=Q97.5)
  ) +
  # ggstance::geom_pointrangeh(
  #   data = tmp1, aes(xmin = Q2.5, xmax = Q97.5),
  #   size = .25, fatten = 1,
  #   position = ggstance::position_dodge2v(.25)
  # ) +
  facet_wrap("Technology") +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none",
    legend.text = element_text(size = 8)
  )
```

```{r}
# Sex
pars %>% 
  filter(str_detect(Parameter, "Sex1:Year:")) %>% 
  mutate(Parameter = "Sex x Year x Tech") %>% 
  ggplot(aes(Estimate, Outcome, shape = Technology)) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(
    "Estimated parameter",
    breaks = scales::pretty_breaks(), 
    expand = expansion(.25)
  ) +
  geom_vline(xintercept = 0, lty = 2, size = .25) +
  geom_linerangeh(
    aes(xmin = Q2.5, xmax = Q97.5), size = .25,
    position = position_dodge2v(.4), show.legend = FALSE
  ) +
  geom_linerangeh(
    aes(xmin = Q10, xmax = Q90), size = .75,
    position = position_dodge2v(.4), show.legend = FALSE
  ) +
  geom_point(
    size = 2, position = position_dodge2v(.4)
  ) +
  facet_wrap("Parameter", scales = "free_x", nrow = 1) +
  theme(
    legend.position = "bottom",
    axis.title.y = element_blank(), 
    panel.spacing.x = unit(12, "pt")
  )
```

```{r, eval = TRUE, warning = FALSE}
options(width = 120)
library(sessioninfo)
session_info()
```
