# GAM

```{r setup, include=FALSE}
pacman::p_load(
  knitr, 
  brms, 
  future,
  scales,
  ggstance,
  patchwork,
  lavaan,
  mgcv,
  tidyverse
)
opts_chunk$set(
  cache = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.retina = 2, 
  fig.width = 6,
  fig.height = 4,
  fig.align = 'center',
  # Save more figure formats
  dev = c("png", "pdf", "tiff"),
  fig.ext = c("png", "pdf", "tiff")
)
theme_set(
  theme_linedraw(base_size = 8) +
    theme(
      panel.grid = element_blank()
    )
)
dir.create("models")
```

## Data

```{r data-load}
mtf <- read_rds("data/mtf.rds")
us <- read_rds("data/us.rds")
yrbs <- read_rds("data/yrbs.rds")

us <- mutate(us, across(sdqc:sdqv, ordered))
yrbs <- mutate(yrbs, across(sad_lonely:suicide_3, ordered))

# Drop rows where all outcome items are missing
# And remove mean outcome (will use latent factors)
mtf <- drop_na(mtf, Depression) %>% select(-Depression)
us <- drop_na(us, Emotion, Conduct) %>% select(-Emotion, -Conduct)
yrbs <- drop_na(yrbs, Suicide) %>% select(-Suicide)
```

## Models

Roadmap:

- For each outcome in each dataset, a multiple indicator latent variable is created with lavaan
- Those latent variables are extracted back to the main data frames
- Latent variables are treated as outcomes in GAMs
- GAMs treat predictors with smooths to allow wiggliness
- Two models are compared: one with smooth X and year, one with smooth X, year, and smooth X by year interaction
- Present all model comparisons and R squareds

## Create latent variables for outcomes

```{r}
fit_sem <- function(data, items, y, missing = "ml") {
  
  # Combine names of items to a string for lavaan model
  items_all <- paste0(items, collapse = " + ")
  
  # Model strings
  sem0 <- str_glue("{y} =~ {items_all}")
  
  ml0 <- sem(sem0, data = data, missing = missing)
  
  return(bind_cols(data, as.data.frame(predict(ml0))))
  
}
```

```{r lavaan-mtf, cache = TRUE}
mtf_mod <- '
  Depression =~ D_B_1 + D_B_2 + D_B_3 + D_B_4 + D_B_5 + D_B_6
'
out <- cfa(mtf_mod, data = mtf, missing = "ml", std.lv = TRUE)
mtf <- bind_cols(mtf, as.data.frame(predict(out)))
mtf$Depression <- as.numeric(scale(mtf$Depression))
```

```{r lavaan-us, cache = TRUE}
us_mod <- '
  Conduct =~ sdqe + sdqg + sdql + sdqr + sdqv
  Emotion =~ sdqc + sdqh + sdqm + sdqp + sdqx 
'
out <- cfa(us_mod, data = us, missing = "pairwise", std.lv = TRUE)
us <- bind_cols(us, as.data.frame(predict(out)))
us$Conduct <- as.numeric(scale(us$Conduct))
us$Emotion <- as.numeric(scale(us$Emotion))
```

```{r lavaan-yrbs, cache = TRUE}
yrbs_mod <- '
  Suicide =~ sad_lonely + suicide_1 + suicide_2 + suicide_3
'
out <- cfa(yrbs_mod, data = yrbs, missing = "pairwise", std.lv = TRUE)
yrbs <- bind_cols(yrbs, as.data.frame(predict(out)))
yrbs$Suicide <- as.numeric(scale(yrbs$Suicide))
```

## Estimate GAMs

```{r}
fit <- function(data, x, y, name) {
  
  # Center year
  data <- mutate(data, Year = Year - 2017)

  # Drop rows with missing predictor
  data <- drop_na(data, all_of(x))
  
  # Ensure max knots
  k_x <- length(table(data[[x]]))
  k_yr <- length(table(data[["Year"]]))

  # Model strings
  model0 <- str_glue("{y} ~ ti(Year, k = {k_yr})")
  model1 <- str_glue("{y} ~ ti(Year, k = {k_yr}) + ti({x}, k = {k_x})")
  model2 <- str_glue(
    "{y} ~ ti(Year, k = {k_yr}) + ti({x}, k = {k_x}) + ti({x}, Year, k = c({k_x}, {k_yr}))"
  )
  
  fit0 <- gam(as.formula(model0), data = data, method = "REML")
  fit1 <- gam(as.formula(model1), data = data, method = "REML")
  fit2 <- gam(as.formula(model2), data = data, method = "REML")
  
  return(
    tibble(
      name, x, y, 
      fit0 = list(fit0), fit1 = list(fit1), fit2 = list(fit2)
    )
  )
}
```

```{r fit-gams}
if (!file.exists("models/gams.rds")) {
  x1 <- fit(mtf, "SM", "Depression", "MTF")
  x2 <- fit(mtf, "TV", "Depression", "MTF")
  x3 <- fit(us, "SM", "Conduct", "US")
  x4 <- fit(us, "TV", "Conduct", "US")
  x5 <- fit(us, "SM", "Emotion", "US")
  x6 <- fit(us, "TV", "Emotion", "US")
  x7 <- fit(yrbs, "DV", "Suicide", "YRBS")
  x8 <- fit(yrbs, "TV", "Suicide", "YRBS")
  
  # Rename variables
  fits <- bind_rows(x1,x2,x3,x4,x5,x6,x7,x8)
  fits <- fits %>% 
    rename(Technology = x, Outcome = y, Study = name) %>% 
    mutate(
      Technology = ifelse(
        Technology %in% c("SM", "DV"), 
        "Social media / device", 
        "Television"
      )
    )
  saveRDS(fits, "models/gams.rds")
} else {fits <- readRDS("models/gams.rds")}
```

```{r}
out <- fits %>% 
  pivot_longer(fit0:fit2) %>% 
  mutate(
    rsq = map_dbl(value, ~summary(.)$r.sq),
    dev = map_dbl(value, ~summary(.)$dev.expl),
    aic = map_dbl(value, AIC)
  ) %>% 
  select(-value) %>% 
  pivot_wider(values_from = rsq:aic)
kable(out, digits = 2)
```

Figure of fitted X - Y lines per year

```{r}
fits
foo <- function(row) {
  fit <- fits$fit2[[row]]
  newx <- distinct(fit[["model"]][,2:3])
  preds <- cbind(newx, predict(fit, newdata = newx, se.fit = TRUE)) %>% 
    as_tibble()
  names(preds)[2] <- "x"
  preds$Year <- preds$Year + 2017
  return(preds)
}
bar <- function(row) {
  fit <- fits$fit2[[row]]
  newx <- as_tibble(fit[["model"]])
  names(newx) <- c("y", "Year", "x")
  newx$Year <- newx$Year + 2017
  return(newx)
}

fits$outs <- map(1:8, foo)
fits$dats <- map(1:8, bar)
p01 <- fits %>% 
  select(-c(fit0:fit2, dats)) %>% 
  unnest(outs) %>% 
  # mutate(Year2 = Year, Year = factor(ifelse(Year < 2005, "< 2005", Year))) %>% 
  ggplot(aes(x, fit, col = Year, group = Year, fill = Year)) +
  scale_colour_viridis_c(
    aesthetics = c("color", "fill"), 
    direction = -1, end = .95, values = c(.6, 1), na.value = "#C9DF2F",
    guide = "legend", 
    breaks = c(2005, 2008, 2011, 2014, 2017), 
    labels = c("2005\n(or before)", 2008, 2011, 2014, 2017)
    ) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  # Can add points but this renders trends invisibly small
  # geom_point(
  #   data = unnest(select(fits, -c(fit0:fit2, outs)), dats),
  #   aes(x=x, y=y), size = .1, alpha = .075,
  #   position = position_jitter(.25, 0, seed = 1)
  # ) +
  geom_ribbon(
    aes(ymin = fit-se.fit*2, ymax = fit+se.fit*2), 
    alpha = .15, col = NA, show.legend = FALSE
  ) +
  geom_line(size = .6) +
  guides(
    color = guide_legend(
      override.aes = list(size = 1.25)
    )
  ) + 
  labs(x = "Technology use", y = "Standardized latent factor score") +
  facet_grid(
    Technology~Outcome, scales = "free_x", labeller = label_wrap_gen(13)
  )
p01
```

```{r Figure3, fig.height = 4.3, fig.width = 6.4}
p1 <- out %>% 
  ggplot(aes(Technology, dev_fit1, shape = Outcome)) +
  scale_y_continuous(
    "Proportion\nvariance explained",
    limits = c(0, .04), 
    labels = function(x) percent(x, 1)
  ) +
  scale_x_discrete(labels = function(x) str_replace(x, "/", "\n/")) +
  geom_point(position = position_jitter(.1, seed = 1), size = 2) +
  scale_shape_manual(values = c(15, 16, 21, 22)) +
  theme(
    legend.position = "none", 
    aspect.ratio = 1, 
    axis.title.x = element_blank()
  )
p2 <- out %>%   
  ggplot(aes(dev_fit2-dev_fit1, aic_fit1-aic_fit2, shape = Outcome)) +
  geom_hline(yintercept = c(-3, 3), lty = 2, size = .25) +
  scale_shape_manual(values = c(15, 16, 21, 22)) +
  geom_point(size = 2) +
  labs(y = "AIC difference") +
  coord_cartesian(ylim = c(-10, 10)) +
  scale_x_continuous(
    "Additional variance explained\nby model with interaction",
    labels = function(x) percent(x, .01)
  ) +
  facet_wrap("Technology", labeller = label_wrap_gen(13)) +
  theme()
p01 / ((p1 | p2) + plot_layout(widths = c(4, 6))) + plot_layout(heights = c(65, 35))
filter(out, Study=="MTF") %>% 
  mutate(delta_aic = aic_fit1-aic_fit2, delta_dev = percent(dev_fit2-dev_fit1)) %>% 
  select(Study:Outcome, starts_with("delta")) 
```

## Session information

```{r, eval = TRUE, warning = FALSE}
options(width = 120)
library(sessioninfo)
session_info()
```
